---
title: "linefollow_analysis"
author: "Nils Wendel Heinrich"
date: "2024-07-17"
output: pdf_document
---

```{r global, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}

library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(latex2exp)
library(ggplot2)
library(fitdistrplus)
library(simr)
library(arrow)
library(MASS)
library(MuMIn)

set.seed(36)
N_iterations <- 100

```


```{r data, echo=FALSE}

trial_data <- read_csv("data/trialwise_data.csv")

# factorization
trial_data$feedback <- factor(trial_data$feedback, levels = c("neutral", "positive", "negative"))

trial_data$input_noise_magnitude <- factor(trial_data$input_noise_magnitude)

```

# H1: Perceived control over spaceship trajectory is expected to be higher for low-input noise spaceship than high-input noise spaceship.

Distributional analysis:
```{r h1_boxcox, echo=FALSE}

lambda_soc <- boxcox(lm(trial_data$SoC ~ 1))

lambda_soc$x[which(lambda_soc$y == max(lambda_soc$y))]

```
referring to:
https://www.statisticshowto.com/probability-and-statistics/normal-distributions/box-cox-transformation/

lambda, the expected value is closer to 0.5 than 1 (1 would be ideal, suggesting no transformation). 0.5 implies a √ transformation, which we will apply. 
"a boxcox distributional analysis implied a square root transformation of the predicted variable"


### null model
```{r h1_lm, echo=FALSE}
soc_null <- lmer(sqrt(SoC) ~ 1 + (1|code), data = trial_data)
```


...
```{r h1_lm, echo=FALSE}

h1_lm <- lmer(sqrt(SoC) ~ input_noise_magnitude + (1|code), data = trial_data)

summary(h1_lm)

```
We detect a significant effect for input_noise_magnitude on SoC (p<.001).

proportional variance explained by our random effect (participant) code
```{r h1_icc, echo=FALSE}

#Intraclass correlation coefficient
performance::icc(h1_lm)

```

```{r h1_r_squared, echo=FALSE}

r_full <- r.squaredGLMM(h1_lm)
# R2m is variance of fixed effects

r_null <- r.squaredGLMM(soc_null)

fixed_effects_variance <- r_full[1,"R2m"] - r_null[1,"R2m"]
fixed_effects_variance
```

```{r h1_coefs, echo=FALSE}

#intercept:
fixef(h1_lm)[1]**2
# mean response across all participants (code) when input_noise_magnitude = 0.5

# slope
fixef(h1_lm)[2]**2
# mean difference in response across all participants (code) when input_noise_magnitude jumps from 0.5 to 2.0

```

parametric bootstrap:
```{r h1_bootstrap, echo=FALSE}

confint(h1_lm, nsim=N_iterations, parm=c('input_noise_magnitude2'), method='boot')

```

bounds of the 95% confidence interval were obtained by a parametric bootstrap with {N_iterations} iterations.

reporting our effect:
...increasing magnitude of input noise significantly decreases SoC (beta={fixef(h1_lm)[2]**2}, CI=[-0.5528189, -0.5290939],  p<.001).


# H2: Perceived control is higher during trials with positive compared to neutral and negative feedback (testing illusion of control)

```{r h2_lm, echo=FALSE}

h2_lm <- lmer(sqrt(SoC) ~ feedback + (1|code), data = trial_data)

summary(h2_lm)

```

```{r h2_coefs, echo=FALSE}

#intercept:
fixef(h2_lm)[1]**2
# mean response across all participants (code) when feedback is neutral

# slopes:
fixef(h2_lm)[2]**2
# mean difference in response across all participants (code) when feedback is positive compared to neutral

fixef(h2_lm)[3]**2
# mean difference in response across all participants (code) when feedback is negative compared to neutral

```

parametric bootstrap:
```{r h2_bootstrap, echo=FALSE}

confint(h2_lm, nsim=N_iterations, parm=c('feedbackpositive', 'feedbacknegative'), method='boot')

```

# H3: Effect of feedback on SoC ratings is higher in high input noise blocks (interaction effect feedback x input noise => SoC)

```{r h3_lm, echo=FALSE}

# full model
h3_lm <- lmer(sqrt(SoC) ~ feedback * input_noise_magnitude + (1|code), data = trial_data)

summary(h3_lm)

```

:( no significant interaction effects.

### Changing contrasts so to compare positive and negative feedback:

```{r h3_lm_valence, echo=FALSE}

# new factorization
trial_data$feedback <- factor(trial_data$feedback, levels = c("positive", "negative", "neutral"))

# full model
h3_lm_valence <- lmer(sqrt(SoC) ~ feedback * input_noise_magnitude + (1|code), data = trial_data)

summary(h3_lm_valence)

# old contrasts
trial_data$feedback <- factor(trial_data$feedback, levels = c("neutral", "positive", "negative"))

```

```{r h3_coefs, echo=FALSE}

#intercept:
fixef(h3_lm)[1]**2
# mean response across all participants (code) when feedback is neutral

# slopes:
fixef(h3_lm)[5]**2
# ...

fixef(h3_lm)[6]**2
# ...

```

parametric bootstrap:
```{r h3_bootstrap, echo=FALSE}

confint(h3_lm, nsim=N_iterations, parm=c('feedbackpositive', 'feedbacknegative', 'input_noise_magnitude2', 'feedbackpositive:input_noise_magnitude2', 'feedbacknegative:input_noise_magnitude2'), method='boot')

```

# H4: Performance (line deviation) is higher for low-input noise spaceship than high-input noise spaceship

Distributional analysis:
```{r avg_boxcox, echo=FALSE}

lambda_avg <- boxcox(lm(trial_data$avg_dist_trialwise ~ 1))

lambda_avg$x[which(lambda_avg$y == max(lambda_avg$y))]

```

The expected value, lambda is close to 0, which implies a log transformation.

```{r h3_lm, echo=FALSE}

# full model
h4_lm <- lmer(log(avg_dist_trialwise) ~ input_noise_magnitude + (1|code), data = trial_data)

summary(h4_lm)

```

```{r h4_coefs, echo=FALSE}

#intercept:
exp(fixef(h4_lm)[1])
# ...

# slopes:
exp(fixef(h4_lm)[2])
# ...


```

proportional variance explained by our random effect (participant) code
```{r h4_icc, echo=FALSE}

#Intraclass correlation coefficient
performance::icc(h4_lm)

```

parametric bootstrap:
```{r h4_bootstrap, echo=FALSE}

confint(h4_lm, nsim=N_iterations, parm=c('input_noise_magnitude2'), method='boot')

```

