---
title: "COAF - Participant Descriptives"
author: "Maren Giersiepen"
date: "2024-07-17"
output:
  html_document:
    keep_md: yes
    theme: cosmo
  pdf_document: default
editor_options:
  chunk_output_type: console
---
# Participant Descriptives of the SPACE Experiment. A collaborative study conducted by the COMPAS and AffAct projects part of the priority program The Active Self. In the task underlying the collected data, participants perform a motor adaptation task in which they steer a stimulus down a pre-defined line using button presses on the keyboard and using the keys [Y] and [M].

## In this script, participant descriptieves are evauated

## Date: 23.07.2024**
## Version: 1**

### Install Packages
```{r, install-packages, message=F, warning=F, include=F}
#install.packages("lattice") 
#install.packages("ggplot2")      # for figures
#install.packages("lavaan")
# install.packages("psych")               # for basic calculations
# install.packages("foreign")             # xxx
# install.packages("car")                 # xxx
# install.packages("corpcor")             # for correlations
# install.packages("Hmisc")               # xxx
# install.packages("polycor")             # xxx
# install.packages("apaTables")           # for exporting tables according to APA style
# install.packages("QuantPsyc")           # xxx
# install.packages("effsize")             # xxx
# install.packages("tidyverse")           # for transforming data
# install.packages("readxl")              # for loading and exporting datasets
# install.packages("writexl")             # for loading and exporting datasets
# install.packages("bannerCommenter")     # for section headers
# install.packages("reshape2")            # for reshaping datasets
# install.packages("responsePatterns")    # for check of response patterns
# install.packages("corrplot")            # for correlation tables
# install.packages("gsubfn")              # for subsetting strings
# install.packages("data.table")          # for creating data tables
# install.packages("RColorBrewer")        # for colours in plots
# install.packages("pals")                # for colours in plots
# install.packages("devtools")
# install.packages("usethis")
# install.packages("lavaan")              # required to run SEMs
# install.packages("lavModel")            # required to run SEMs
# install.packages("ROCR")                # required to create ROC curves
# install.packages("janitor")
# install.packages("foreign")             # to load SPSS datafiles
# install.packages("sjlabelled")          # to remove SPSS labels from dataset
# install.packages("mice")
# install.packages("multcomp")
# install.packages("ggpubr")
# install.packages("PairedData")
# install.packages("lme4")
# install.packages("emmeans")
# install.packages("multcompView")
# install.packages("nlme")
# install.packages("ez")
# install.packages("BayesFactor")
# install.packages("svglite")
```

### Load Packages
```{r, load-packages, message=F, warning=F, include=F} 
library("dplyr")
library("ggplot2")
library("ggpubr")
library("lattice")
library("svglite")
library("devtools")
library("gridExtra")
library("readxl")
library("tidyverse")
#library("pals")
#library("lavaan")
#library("psych")
#library("foreign")
#library("car")
#library("corpcor")
#library("Hmisc")
#library("polycor")
#library("apaTables")
#library("QuantPsyc")
#library("effsize")
#library("writexl")
#library("bannerCommenter")
#library("reshape2")
#library("responsePatterns")
#library("corrplot") 
#library("gsubfn")
#library("purrr")
#library("data.table")
#library("RColorBrewer")
#library("usethis")
#library("lavaan")
#library("ROCR")
#library("githubinstall")
#library("stats")
#library("tidyr")
#library("report")
#library("janitor")
#library("foreign")
#library("sjlabelled")
#library("mice")
library("rstatix")
#library("multcomp")
#library("PairedData")
library("lme4")
#library("emmeans")
#library("multcompView")
#library("nlme")
#library("ez")
#library("BayesFactor")
```

### Load Data
```{r, load behavioral data, include = F}
#1) load experimental data of all participants
main_folder <- "H:/Maren/Experiments/COAF/Main/data/experiment_preprocessed"
data_files <- list.files(main_folder, pattern = "data", full.names = TRUE)

final_data <- read.csv(data_files[1])
trialwise_data <- read.csv(data_files[2])

#1) load questionnaire data of all participants
main_folder <- "H:/Maren/Experiments/COAF/Main/data/questionnaires"
data_file <- list.files(main_folder, pattern = "data", full.names = TRUE)

for (file in data_file) {
  questionnaire_raw_allPP <- read_excel(file)
}
```

```{r, set correct variable levels, include = F}
final_data$code <- as.factor(final_data$code)
trialwise_data$code <- as.factor(trialwise_data$code)

final_data$input_noise_magnitude <- as.factor(final_data$input_noise_magnitude)
trialwise_data$input_noise_magnitude <- as.factor(trialwise_data$input_noise_magnitude)

final_data$feedback <- as.factor(final_data$feedback)
trialwise_data$feedback <- as.factor(trialwise_data$feedback)

```

```{r, evaluate participant descriptives, using the entire sample, include = T}

#Sample size
N_total <- length(unique(final_data$code))
print(N_total) 

#Age
#create one row per participant
ppwise_data <- final_data %>%
  group_by(code) %>%
  slice(1)

#mean age
M_age <- mean(ppwise_data$age) 
print(M_age)

#sd_age 
SD_age <- sd(ppwise_data$age) 
print(SD_age)

#range of age values (should be between 18 and 45 according to inclusion criteria)
min_age <- min(ppwise_data$age)
print(min_age) 

max_age <- max(ppwise_data$age)
print(max_age) 

#Gender distribution
females <- sum(ppwise_data$gender == "f")
print(females) 

males <- sum(ppwise_data$gender == "m")
print(males) 


nonbinary <- sum(ppwise_data$gender != "m" & ppwise_data$gender != "f")
print(nonbinary) 
```

```{r, evaluate participant performance, including the entire sample, include = T}


#(1) Below/Above average performance in steering the spaceship

#calculate the average distance of each participant for each input noise condition
descriptives_dist_ppwise <- trialwise_data %>%
  dplyr::group_by(input_noise_magnitude, code) %>%
  dplyr::summarise(m_dist = mean(avg_dist_trialwise), sd_dist = sd(avg_dist_trialwise))
head(descriptives_dist_ppwise)

#identify outliers as defined in identify_outliers() in R:
# 1) Outliers: values < [1st quartile – 1.5 *interquartile range] or > [3rd quartile + 1.5 * interquartile range])
# 2) Extreme Outliers: values < [1st quartile – 3 *interquartile range] or > [3rd quartile + 3 * interquartile range]).  

outliers <- descriptives_dist_ppwise %>% 
  group_by(input_noise_magnitude) %>%
  rstatix::identify_outliers(m_dist)
print(outliers)
#=> two extreme outliers in both low and high input noise conditions, with pp12 and pp14 being the extreme outliers in both cases
  

#for visualization, create a boxplot for each condition and color outlier participants in red (geom_boxplot considers values outside 1.5 x IQR from the lowest and highest data point an outlier)
boxplot_performance_outlier <- ggplot(descriptives_dist_ppwise, aes(x = input_noise_magnitude, y = m_dist)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Boxplot for high and low input noise",
       x = "Input Noise",
       y = "Line Distance") +
  theme_minimal()
print(boxplot_performance_outlier)

#(2) Unrealistic, non-compliant SoC values (e.g., same SoC rating across all trials / trials within conditions)
#create a dataset counting frequency of individual SoC values per participant
SoC_freq_ppwise <- trialwise_data %>%
  dplyr::group_by(code, SoC) %>%
  dplyr::summarise(count = n(), .groups = 'drop')
head(SoC_freq_ppwise)

#plot the data 
#histogram of participant rating frequencies for the entire sample
ggplot(trialwise_data, aes(x = SoC)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()

#as plotting all 50 participants in the same figure makes the plot difficult to read, create subsets of 10 participants to plot in an individual figure
pp_set1 <- trialwise_data %>% filter(code %in% 1:10)
pp_set2 <- trialwise_data %>% filter(code %in% 11:20)
pp_set3 <- trialwise_data %>% filter(code %in% 21:30)
pp_set4 <- trialwise_data %>% filter(code %in% 31:40)
pp_set5 <- trialwise_data %>% filter(code %in% 41:50)

#overview of rating distribution across all participants
ggplot(SoC_freq_ppwise, aes(x = factor(code), y = count, fill = factor(SoC))) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Each SoC Rating per Participant",
       x = "Participant (code)",
       y = "Frequency (count)",
       fill = "SoC") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


#histogram of participant rating frequencies in subsets of the sample 
ggplot(pp_set1, aes(x = SoC)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()

#=> participant 2 as very little variation in SoC ratings

#histogram of participant rating frequencies in subsets of the sample 
ggplot(pp_set2, aes(x = SoC)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()

#=> participant 14 only provided SoC Ratings = 1

#histogram of participant rating frequencies in subsets of the sample 
ggplot(pp_set3, aes(x = SoC)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()

#histogram of participant rating frequencies in subsets of the sample 
ggplot(pp_set4, aes(x = SoC)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()

#histogram of participant rating frequencies in subsets of the sample 
ggplot(pp_set5, aes(x = SoC)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()

#=> Based on the above performed analysis of participants' performance and SoC ratings, participant 12 and 14 are excluded from all further analyses

#finally, get an overview of participant SoC rating distribution for high and los input noise trials
ggplot(pp_set1, aes(x = SoC, fill = input_noise_magnitude)) +
  geom_histogram(binwidth = 1, position = "dodge" , color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()

#finally, get an overview of participant SoC rating distribution for high and los input noise trials
ggplot(pp_set2, aes(x = SoC, fill = input_noise_magnitude)) +
  geom_histogram(binwidth = 1, position = "dodge" , color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()


#finally, get an overview of participant SoC rating distribution for high and los input noise trials
ggplot(pp_set3, aes(x = SoC, fill = input_noise_magnitude)) +
  geom_histogram(binwidth = 1, position = "dodge" , color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()


#finally, get an overview of participant SoC rating distribution for high and los input noise trials
ggplot(pp_set4, aes(x = SoC, fill = input_noise_magnitude)) +
  geom_histogram(binwidth = 1, position = "dodge" , color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()


#finally, get an overview of participant SoC rating distribution for high and los input noise trials
ggplot(pp_set5, aes(x = SoC, fill = input_noise_magnitude)) +
  geom_histogram(binwidth = 1, position = "dodge" , color = "black") +
  facet_wrap(~ code, scales = "free_y") +
  labs(title = "Distribution of SoC Ratings per Participant",
       x = "SoC Rating",
       y = "Frequency") +
  theme_minimal()

```


```{r, evaluate participant performance, including the final sample, include = T}

#exclude pp12 and pp14 from the data
final_ppset <- ppwise_data %>%
  subset(code != 12 & code != 14)

#Sample size
N_total <- length(unique(final_ppset$code))
print(N_total) 

#Age
#mean age
M_age <- mean(final_ppset$age) 
print(M_age)

#sd_age 
SD_age <- sd(final_ppset$age) 
print(SD_age)

#range of age values (should be between 18 and 45 according to inclusion criteria)
min_age <- min(final_ppset$age)
print(min_age) 

max_age <- max(final_ppset$age)
print(max_age) 

#Gender distribution
females <- sum(final_ppset$gender == "f")
print(females) 

males <- sum(final_ppset$gender == "m")
print(males) 

nonbinary <- sum(final_ppset$gender != "m" & final_ppset$gender != "f")
print(nonbinary) 

```

